---
title: " Introductory Data Science for Innovation. W1 - Exercises"
author: "David Eggleton"
date: '2023-09-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the snippets below try the following:

# 1. Do simple computations

Simple computation (Use any number. Try it in the Console too. Remember it's Ctrl + Enter to run a single line in a .R file but in a .Rmd file you can also click 'play').
Try out:

* addition  
* subtraction  
* multiplication  
* division  

```{r basic arithmetic, echo = FALSE}


```

# 2. Use variable 'a' in cumulative steps

* Assign the value 1 to a variable called a. Key command for <- is 'Alt'/'Option' + '-'
* Add 1 to a and assign it back to a
* Subtract a from 10 and assign it back to a
* Multiply a by 4 and assign it back to a
* Divide a by 2 and reassign it back to a
* Create a new variable for the 'expected value of a' and set it equal to 16
* Check if value of a is right (if it returns true you are right, if false you are wrong). 
``` {r a_playtime, echo = FALSE}

```

Print the value of a

```{r print, echo = FALSE}

```

Go back in code and assign a different value to a (such as 10 or 5) and rerun the code above. Just play around - you can't break it and even if you do just revert to the git.

```{r scratch, echo = FALSE}



```

# 3. Including Plots

You can also embed plots so it's a bit easier to see, for example:

```{r plot, echo= FALSE}
data(Titanic)
Titanic <- as.data.frame(Titanic) 

hist(Titanic$Freq)
```

Note that the `echo = FALSE` parameter above was added to the code chunks to prevent printing of the R code that generated the plot. Try changing it to `TRUE` and see what (if anything) happens!

# 4. Work with packages

Try to install a library whether in the console or here. Tidyverse is a good idea because you're going to be using it a lot. Syntax is a bit weird but you'll get used to it. Installation requires quotes, loading the library doesn't
```{r packageInstall, echo = FALSE}



```
